A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
related

\begin{abstract}
  The \eval function turns text into code. From the point of view of
  reasoning about software system, uses of \eval turn the program's code
  from a statically known quantity that can be reasoned about and analyzed
  into one of the inputs to the program. Thus, widespread use of \eval
  hinders our ability to provide safety and security guarantees for
  software. Understanding how \eval is used in practice is key to finding
  ways to mitigate its damage. One question that has not been studied is
  whether the usage of \eval is dependent on the application domain or other
  features of the language. This paper is a large-scale study of a corpus of
  \CorpusAllCodeRnd lines of data science libraries and end-user written in
  the R language. We performed a simple static analysis and more
  sophisticated dynamic analysis to confirm that \eval is indeed in wide
  spread use, and to categorized patters of usage. We found that \eval in
  this code base is more dangerous in some ways, and safer in others, than
  what was previously reported for JavaScript.
\end{abstract}

@@

To understand the behavior of eval we perform a combination of static
analysis, dynamic analysis and manual code inspection. The results presented
here summarize our findings.

@@

The difference between these design choices has an impact on the ability of
compilers to optimize, and of analysis tools to reason about, code. The
impact of a top level \eval is that any global definition may be affected
but the current function can continue to execute unchanged. A local \eval
may change the value of local variables, their types, or even their
presence. This is a much more invasive change for reasoning about code and
may invalidate common optimizations such as constant folding or copy
propagation. The designers of Julia were aware of this difference and chose
to restrict \eval to preserve their ability to compile
code~\cite{oopsla18a}. They even added a versioning mechanism to function
definitions to ensure that new code added during an \eval would not
invalidate optimizations such as inlining.  R's ability to affect any one of
the callers of a function is even more invasive as the impact of \eval is
non-local. Such unfettered power is difficult to deal with.

The intuition behind the soundiness manifesto~\cite{soundy} is that while
the use of a single \eval may invalidate every single invariant computed by
a static analysis or optimizer, this worst case usually does not happen.  In
many cases, \eval is used carefully and has limited effect. This is a
hypothesis that must be validated and, to be useful, we should have the
means to provide some reasonable guesses as to which side-effects will
result from any particular \eval call.

Previous work looked at eval in JavaScript

Othere languages

Dynamic analysis

@@

@@

\begin{figure*}[!tb]\centering\includegraphics[width=\linewidth]
{pipeline.pdf}\caption{Pipeline}label{fig:pipeline}
\end{figure*}

@@

\begin{figure*}[!tb]\centering\includegraphics[width=\linewidth]
  {corpus.pdf}\caption{CRAN packages}\label{fig:corpus}
\end{figure*}

>>>>

\begin{abstract}
  The \eval function turns text into code. From the point of view of
  reasoning about software system, uses of \eval turn the program's code
  from a statically known quantity that can be reasoned about and analyzed
  into one of the inputs to the program. Thus, widespread use of \eval
  hinders our ability to provide safety and security guarantees for
  software. Understanding how \eval is used in practice is key to finding
  ways to mitigate its damage. One question that has not been studied is
  whether the usage of \eval is dependent on the application domain or other
  features of the language. This paper is a large-scale study of a corpus of
  \CorpusAllCodeRnd lines of data science libraries and end-user written in
  the R language. We performed a sophisticated dynamic analysis to confirm
  that \eval is indeed in wide spread use, and to categorized patters of
  usage. We found that \eval in this code base is more dangerous in some
  ways, and safer in others, than what was previously reported for
  JavaScript.
\end{abstract}

@@

To understand the behavior of eval we perform a dynamic analysis and manual
code inspection. The results presented here summarize our findings.

@@

The difference between these design choices has an impact on the ability of
compilers to optimize, and of analysis tools to reason about code. The
impact of a top level \eval is that any global variable may be modified, but
currently running functions can continue to execute unchanged. A local
\eval, may change the values of local variables, their types, or even their
presence. This is a much more invasive change, one that may invalidate
currently executing functions as it may change the results of common
optimizations such as constant folding or copy propagation. The designers of
the Julia programming language were aware of this difference and
purposefully chose to restrict \eval to the top level in order to preserve
their ability to optimize code~\cite{oopsla18a}. They even added a
versioning mechanism to function definitions to ensure that code added
during an \eval would not invalidate optimizations such as inlining.  R's
ability to affect any one of the callers of a function is even more invasive
as the impact of \eval is non-local. Such unfettered power is difficult to
deal with.

The intuition behind the soundiness manifesto~\cite{soundy} is that while
the use of a single \eval may invalidate every single invariant computed by
a static analyzer, this worst case usually does not happen.  In many cases,
\eval is used carefully to limited effect. This hypothesis must be validated
and, for it to be useful, we need to have a way to make some educated
guesses as to which side-effects will result from any particular \eval.

Richards et al.~\cite{ecoop11} provided the first large-scale study of the
runtime behavior of \eval in JavaScript. They analyzed a corpus of the
10,000 most popular websites. They used an instrumented web browser to
gather execution traces and perform a dynamic analysis.  Their results show
that \eval is pervasive with 82\% of the most popular websites using it. The
reasons for using \eval include the desire to load code on demand (to reduce
the time it takes to display a page), the deserialization of JSON data
(about 37\% of all calls) and lightweight meta-programming to customize web
pages.  While many uses \eval were legitimate, many were unnecessary and
could be replaced with equivalent and safer code.  The reported on a
categorization of inputs to \eval that covers the vast majority of input
strings.  Restricting themselves to \eval in which all named variables refer
to the global scope, many patterns can be replaced by more disciplined
code~\cite{oopsla12b}.  The limitation of that work are that there was no
measure of code coverage, so the numbers presented can be a lower bound on
the possible behaviors.  Furthermore in 2011, JavaScript usage is quite
different from today. Their work captured very few libraries and none of the
more recent Node.js ecosystem.

@@

\begin{figure*}[!tb]\centering\includegraphics[width=.8\linewidth]
{pipeline.pdf}\caption{Analysis Pipeline}\label{fig:pipeline}
\end{figure*}

@@

@@

\begin{figure*}[!tb]\centering\includegraphics[width=.8\linewidth]
  {corpus.pdf}\caption{CRAN packages}\label{fig:corpus}
\end{figure*}
